Precise Data Flow Analysis in the Presence of Correlated Method Calls

### 2014-12-13 content/intro.tex
eliminate TODOs in introduction; improve formatting of title

We present a formalization of the transformation and prove its correctness
as follows. 
First, we derive an IDE problem that is equivalent to the original IFDS problem,
as described by Reps et al.~\cite{sagiv1996precise}.
Solving this ``Equivalence-IDE Problem'' with the IDE algorithm yields a result that
is equivalent to the one obtained by applying the IFDS algorithm to the original IFDS problem.
Then, we derive a ``Correlated Calls 
IDE Problem'' from the original IFDS problem and show 
that the solution to this problem, when converted to an IFDS result, is more precise
than the original IFDS result.
We also show that the correlated-calls analysis is sound, i.e., that it never considers concrete execution paths as infeasible.
This is illustrated schematically
in Figure~\ref{fig:transformations}%
\footnote{ 
  The labels $\resultIDE$, $\resultIFDS$, $\backEq$, $\backCC$, $\transEq$, and $\transCC$
  on edges in 
  Figure~\ref{fig:transformations} reflect a number of mappings and projections 
  that will be defined in Section~\TODO{ref} and that can be ignored here.
}.\mtodo{In the footnote, should we just say "the labels" since there are so many?}

@@

The remainder of this paper is organized as follows.
%
Section~\ref{sec:MotivatingExample} presents a motivating example.
%
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
%
Section~\ref{sec:ccdatastr} presents the correlated-calls transformation
and a proof of its correctness.
%
The implementation of our approach and preliminary experiments are discussed
in Section~\ref{chapter:eval}.
%
Related work is discussed in Section~\TODO{ref}.
%
Finally, conclusions and directions for future work are presented in Section~\TODO{ref}.

\commentout{
Static program analysis aims to discover properties of computer programs without running them.
Static analysis has applications in compiler optimization, development of programming tools, and computer security, among others.
As an example, we might want to analyze a program to know which variables are constants. 
We could then write a compiler optimization that ensures that the values of those variables are computed only once.
Alternatively, we could use the information about constant variables in an integrated development environment; for instance, to notify the user when an if-expression executes only one of its branches because its test condition has a constant value.

There are demonstrable limits on what information we can obtain about a program without running it.
Rice's theorem states that verifying any non-trivial property of a program is an undecidable problem~\cite{rice1953classes}. However, it is sometimes possible to design an algorithm that \textit{over}- or \textit{underapproximates} the solution that we are seeking.

\textit{Data-flow analysis} is an area of program analysis whose goal is to compute approximations of certain information (for example, which variables must be constants) for each program point.

Other examples of data-flow analyses are \textit{reaching definitions} (finding out up to which instruction a given assignment of a variable must be valid) and \textit{available expressions} (retrieving the expressions in the program that do not need to be recomputed at a given program point).

Another example of a data-flow analysis is \textit{taint analysis}~\cite{tripp2009taj}. Taint analysis discovers if ``secret'' values, like passwords or other confidential user information, can leak to an external observer. Methods that generate secret values, e.g. those that read user input, are called \textit{sources}. Methods that can leak information, e.g. those that write data to a file or send data through a network, are called \textit{sinks}. The goal of taint analysis is to find out whether data can propagate from sources to sinks.

An important property of a data-flow analysis is \textit{precision}.
Precision reflects how closely a data-flow-analysis result over- or underapproximates the information we are interested in. In the case of taint analysis, let $T$ be the number of sinks that the analysis considers to leak secret information, and $R$ the real number of potential information leaks. The smaller the difference between $T$ and $R$, the greater the precision of the taint analysis.

Data-flow analyses operate on \textit{control-flow graphs} that model the order in which the instructions of a program are executed.
A data-flow-analysis problem defines \textit{flow functions} that represent how data is propagated along the edges of the control-flow graph. The \textit{confluence operator} specifies how the data that has been computed along different paths should be merged when the paths join.

Since a control-flow graph is an overapproximation of the possible flows of control in concrete executions of a program, the graph may contain \textit{infeasible} paths that cannot occur at runtime.

One way to improve the precision of a data-flow analysis is to detect and eliminate infeasible paths.

Our goal is to improve the precision of solutions to problems that can be solved by the \textit{Inter-procedural Finite Distributive Subset} (IFDS) algorithm~\cite{reps1995precise}.
The IFDS algorithm is a general data-flow algorithm that can compute solutions to various data-flow problems, like reaching definitions, available expressions, and taint analysis.

We improve the precision of IFDS problem solutions by eliminating infeasible paths that occur in object-oriented programs in the presence of \textit{correlated method calls}~--- polymorphic calls that are invoked on the same object~\cite{DBLP:journals/scp/Tip15}.
}

>>>>

We present a formalization of the transformation and prove its correctness
as follows. 
First, we derive an IDE problem that is equivalent to the original IFDS problem,
as described by Reps et al.~\cite{sagiv1996precise}.
Solving this ``Equivalence-IDE Problem'' with the IDE algorithm yields a result that
is equivalent to the one obtained by applying the IFDS algorithm to the original IFDS problem.
Then, we derive a ``Correlated Calls 
IDE Problem'' from the original IFDS problem and show 
that the solution to this problem, when converted to an IFDS result, is more precise
than the original IFDS result.
We also show that the correlated-calls analysis is sound, i.e., that it never considers concrete execution paths as infeasible.
This is illustrated schematically
in Figure~\ref{fig:transformations}%
\footnote{ 
  The labels %$\resultIDE$, $\resultIFDS$, $\backEq$, $\backCC$, $\transEq$, and $\transCC$
  on edges in 
  Figure~\ref{fig:transformations} reflect a number of mappings and projections 
  that will be defined in Section~\ref{sec:ccdatastr} and that can be ignored here.
}.

@@

The remainder of this paper is organized as follows.
%
Section~\ref{sec:MotivatingExample} presents a motivating example.
%
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
%
Section~\ref{sec:ccdatastr} presents the correlated-calls transformation
and a proof of its correctness.
%
The implementation of our approach and preliminary experiments are discussed
in Section~\ref{chapter:eval}.
%
Related work is discussed in Section~\ref{sec:Related}.
%
Finally, conclusions and directions for future work are presented in Section~\ref{sec:Conclusions}.
